Quantifying the importance of an inference model in Bayesian phylogenetics

### 2020-03-23 content.tex
3248 words

The goal of \verb;pirouette; is to quantify the importance of a tree prior.
It does so by measuring the inference error made 
for a given reconstructed phylogeny, 
simulated under a (usually novel) diversification model.
We refer to the true model that generated the given tree
as the 'generative tree model' $\mathit{p_{G}}$.
We define the birth-death (BD) model (\cite{nee1994reconstructed}) as
the standard tree model, as many (non-standard) tree models 
have a parameter setting for which they reduce to this model. 
For example, a protracted birth-death (PBD) model [\cite{etienne2012prolonging}] 
reduces to a BD model when the speciation-completion rate 
is infinite [i.e. rate $\lambda$ in \cite{etienne2014estimating}, eqs.~(2b) and (2c)].
When benchmarking a novel tree model, 
one will typically construct phylogenies 
for different combinations of the diversification model's parameters, 
to assess under which scenarios the inference error cannot be neglected. 
While we recommend many replicate simulations 
when assessing a novel tree prior, 
our examples contain only one replicate,
as the goal is to show the workings of \verb;pirouette;,
instead of doing an extensive analysis.
The supplementary materials, however, do show the results
of replicated runs under multiple settings.

@@

We show the usage of \verb;pirouette; on an tree generated 
by the non-standard diversity-dependent tree model \cite{DDD, etienne2011diversity},
which is a birth-death model with a speciation rate that is dependent on the number of species.

We use a small tree with six taxa, to keep the figure more readable,
with an arbitrary crown age of ten time units.

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_30/example_30_314/true_tree.png}
  \caption{
    The tree derived from a DD process, 
    as created by listing~\ref{lst:create_dd_tree}.
  }
\end{figure}

>>>>

The goal of \verb;pirouette; is to quantify the importance of a tree prior.
It does so by measuring the inference error made 
for a given reconstructed phylogeny, 
simulated under a (usually novel) diversification model.
We refer to the true model that generated the given tree
as the 'generative tree model' $\mathit{p_{G}}$.
We define the birth-death (BD) model (\cite{nee1994reconstructed}) as
the standard tree model, as many (non-standard) tree models 
have a parameter setting for which they reduce to this model. 
One such example is the diversity-dependent (DD) tree model (\cite{DDD, etienne2011diversity}):
the DD model is a BD model with a speciation rate that is dependent 
on the number of species and a carrying capacity. For a carrying capacity
of infinity, the DD model reduces to a BD model.
When benchmarking a novel tree model, 
one will typically construct phylogenies 
for different combinations of the diversification model's parameters, 
to assess under which scenarios the inference error cannot be neglected. 
While we recommend many replicate simulations 
when assessing a novel tree prior, 
our examples contain only one replicate,
as the goal is to show the workings of \verb;pirouette;,
instead of doing an extensive analysis.
The supplementary materials, however, do show the results
of replicated runs under multiple settings.

@@

We show the usage of \verb;pirouette; on an tree generated 
by the non-standard diversity-dependent (DD) tree model \cite{DDD, etienne2011diversity},
which is a BD model with a speciation rate that is dependent on the number of species.

We use the tree as shown in figure \ref{fig:dd_tree}.
The with six taxa, to keep the figure more readable,
with an arbitrary crown age of ten time units.

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_30/example_30_314/true_tree.png}
  \caption{
    The tree derived from a DD process.
  }
  \label{fig:dd_tree}
\end{figure}
